---
title: "236523 - Introduction to Bioinformatics - Spring 21 - HW1"
output:
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Roy Bernea - 316315829 ; Alon Hacohen - 311587653

# Qusetion 1

Following is the code and answers for the questions under the "Question 1" Section - as instructed in the assignment:

### Q1.1
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

x1<-runif(6, min = 1, max = 20)
x2<-runif(6, min = 1, max = 20)
x3<-runif(6, min = 1, max = 20)
x4<-runif(6, min = 1, max = 20)
mat1<-rbind(x1,x2,x3,x4)
print(mat1)

```

### Q1.2
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

sum_of_rows<-apply(mat1,1,sum)
print(sum_of_rows)

```

### Q1.3
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

tentative.normalization <- function(mat,npar){
  max_vector_row <-apply(mat,1,max)
  max_vector_col <-apply(mat,2,max)
  if(npar==1){
    return(mat/max_vector_row)
  }
  if(npar==2){
    return(t(t(mat)/max_vector_col))
  }
  return(mat)
}

```

### Q1.4
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

mat1.norm.rows<-tentative.normalization(mat1,1)
mat1.norm.columns<-tentative.normalization(mat1,2)

print(mat1.norm.rows)
print(mat1.norm.columns)

```

### Q1.5
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

heatmap(mat1.norm.rows)
heatmap(mat1.norm.columns)

```


# Question 2

Following is the code and answers for the questions under the "Question 2" Section - as instructed in the assignment:

### Q2.1
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

data1.file <- file.path("./data1.csv")
data2.file <- file.path("./data2.csv")
data1 <- read.csv(data1.file)
data2 <- read.csv(data2.file)

```

### Q2.2
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

data1_type <- class(data1)
print(data1_type)

```
The type of data1 is **data.frame**.
As you may see in our code, the function used for this is the class() function, which allows us to figure out the data structure of an object.

### Q2.3
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

library(dplyr)

data2_wo_NA <- filter(na.omit(data2))
control <- filter(data2_wo_NA, Treatment == "Placebo")
treatment <- filter(data2_wo_NA, Treatment == "HF123")
print(control)
print(treatment)

```
* We chose to already remove the "NA" rows from the data, as they do not provide any additional information for our uses. 

### Q2.4
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

control_vec_BSL <- apply(control["Blood_Sugar_Levels"],1,as.double)
treatment_vec_BSL <- apply(treatment["Blood_Sugar_Levels"],1,as.double)
mean_control <- mean(control_vec_BSL)
mean_treatment <- mean(treatment_vec_BSL)
mean_difference <- abs(mean_control-mean_treatment)

print(mean_difference)

```
As you may notice, the Mean Difference is **3.041439**.
This mean difference is **not significant** - this variable is not of any statistical value regarding the context of the question, and does not really help us determine any conclusions about the data.

### Q.2.5
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

hist(control_vec_BSL, col=rgb(0,0,1,0.3), main="Blood Sugar Levels, Control & Treatment groups", xlab="Blood Sugar Level", xlim=c(15,40))
hist(treatment_vec_BSL, col=rgb(1,0,0,0.5), add=T)

```
* Please notice that the *blue* represents the **control** group, and the *red* represents the **treatment** group.

### Q.2.6
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

boxplot(control_vec_BSL, treatment_vec_BSL, names=c("Control","Treatment"), main="Control & Treatment groups - Boxplot (data2)")

```

### Q2.7
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

t.test(control_vec_BSL, treatment_vec_BSL)

```
The p-value calculated in the t-test is **0.05894**.
This p-value is *larger than 0.05*, which means that we can not contradict the **Null Hypothesis** based on it.
In simpler words, it means that we can not state that the difference between the *control* group and the *treatment* group is significant (at least not based on *data2*), therefore we are unsure if the treatment has any real effect on the patients.

### Q2.8
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

data1_wo_NA <- filter(na.omit(data1))
control1 <- filter(data1_wo_NA, Treatment == "Placebo")
treatment1 <- filter(data1_wo_NA, Treatment == "HF123")
control_vec_BSL_1 <- apply(control1["Blood_Sugar_Levels"],1,as.double)
treatment_vec_BSL_1 <- apply(treatment1["Blood_Sugar_Levels"],1,as.double)

boxplot(control_vec_BSL_1, treatment_vec_BSL_1, names=c("Control","Treatment"), main="Control & Treatment groups - Boxplot (data1)")

t.test(control_vec_BSL_1, treatment_vec_BSL_1)

```
This time, based on data1, the p-value calculated in the t-test is **1.737x10^-12)^**.
This p-value is *smaller than 0.05*, which means that we **can** indeed contradict the **Null Hypothesis** based on it.
In simpler words, it means that this time, based on data 1, we **can** state that the difference between the *control* group and the *treatment* group **is significant**, therefore it seems that the treatment does have a real effect on the patients.

### Q2.9
When comparing between the t-tests of data1 vs t-test of data2, we trust the result of the first test more (meaning the **t-test performed on data1**).
The reason for that is that data1 includes 846 records, while data2 includes only 24 records.
This means we rely on data from a lot more patients when performing the t-test on data1, which allows the distribution to probably be closer to the one in real-life. This also affects the p-value of the t-test, as seen in previous questions.
Another way to see that data1 is more diverse than data 2 (which will lead to better results) is looking at the *Gender* column - examining the records of data1 and data2 shows us that the majority of the patients in data2 (over 90%) are females, while the patients in data1 divide into almost even groups of males and females.


# Question 3

Following is the code for generating the GRanges objects - as instructed in the assignment:

### Q3.1
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

CLIP_data <- read.table(file.path("./CLIP_data.BED"),header=TRUE)
library("GenomicRanges")
gr <- GRanges(CLIP_data)
gr

```

### Q3.2
```{r,echo=TRUE,eval=TRUE, warning=FALSE, message=FALSE}

gr_flank <- flank(gr, width=200, start=FALSE, both=FALSE)
gr_flank

```

