---
title: "HW1_Markdown"
#date: "4/20/2021"
output:
  word_document: default
editor_options: 
  chunk_output_type: console
---



*** Setup ***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



*** Roy's Parts - Q1 ***

## FIX THIS

## Q1.2

```{r Q1.1}
tentative.normalization <- function(mat,npar){
  max_vector_row <-apply(mat,1,max)
  max_vector_col <-apply(mat,2,max)
  if(npar==1){
    return(mat/max_vector_row)
  }
  if(npar==2){
    return(t(t(mat)/max_vector_col))
  }
  
  return(mat)
}
x1<-runif(6, min = 1, max = 20)
x2<-runif(6, min = 1, max = 20)
x3<-runif(6, min = 1, max = 20)
x4<-runif(6, min = 1, max = 20)
mat1<-rbind(x1,x2,x3,x4)
sum_of_rows<-apply(mat1,1,sum)
mat1.norm.rows<-tentative.normalization(mat1,1)
mat1.norm.columns<-tentative.normalization(mat1,2)
heatmap(mat1.norm.rows)
heatmap(mat1.norm.columns)
```


*** Q2 ***


```{r Q2}

## Q2.1
data1.file <- file.path("./data1.csv")
data2.file <- file.path("./data2.csv")
data1 <- read.csv(data1.file)
data2 <- read.csv(data2.file)

## Q2.2
data1_type <- typeof(data1)
print(data1_type)

## Q2.3
data2_wo_NA <- filter(na.omit(data2))
control <- filter(data2_wo_NA, Treatment == "Placebo")
treatment <- filter(data2_wo_NA, Treatment == "HF123")

## Q2.4
print(control)
control_vec_BSL <- 
  apply(control["Blood_Sugar_Levels"],1,as.double)
treatment_vec_BSL <- 
  apply(treatment["Blood_Sugar_Levels"],1,as.double)
mean_control <- mean(control_vec_BSL)
mean_treatment <- mean(treatment_vec_BSL)
mean_difference <- mean_control - mean_treatment
print(mean_difference)

## Q2.5
library(ggplot2)
library(RColorBrewer) # import ggplot2

# DO THE HIST
hist(control_vec_BSL, col=rgb(0,0,1,0.3), main="histogram of blood sugar levels in the control and treatment groups", xlab="Blood sugar level")
hist(treatment_vec_BSL, col=rgb(1,0,0,0.5), add=T)

# to check parameters


## Q2.6
# Do The Boxplot

## Q2.7.1
t.test(control_vec_BSL, treatment_vec_BSL)

## Q2.7.2
data1_wo_NA <- filter(na.omit(data1))
control1 <- filter(data1_wo_NA, Treatment == "Placebo")
treatment1 <- filter(data1_wo_NA, Treatment == "HF123")

control_vec_BSL_1 <- 
  apply(control1["Blood_Sugar_Levels"],1,as.double)
treatment_vec_BSL_1 <- 
  apply(treatment1["Blood_Sugar_Levels"],1,as.double)
mean_control_1 <- mean(control_vec_BSL_1)
mean_treatment_1 <- mean(treatment_vec_BSL_1)
mean_difference_1 <- mean_control_1 - mean_treatment_1
print(mean_difference_1)

t.test(control_vec_BSL_1, treatment_vec_BSL_1)



## Q2.8




```

Q.2.2 The type of data1 is *list*.

Q.2.4
Is this mean difference?
# DO THIS

Q2.7
p-value > 0.05 => does not condradict the null hypothesis, therefore we cannot say it is not noise

Q2.8
p-value < 0.05 => does indeed contradict the null hypothesis, therefore we can say it is not noise

Q2.9
We trust data1 more because its bigger





*** Alon's Parts - Q3 ***








ggplot(data = data2_wo_NA,mapping =  aes(x=Treatment, y=Blood_Sugar_Levels, color=Placebo)) +
    geom_point()


ggplot(data = data2_wo_NA,mapping =  aes(x=Treatment, y=Blood_Sugar_Levels)) +
    geom_boxplot()


hist(x = control_vec_BSL,treatment_vec_BSL)
